---
title: "Biomass_depth analysis"
author: "Amieroh, Kim and Yameen"
date: "20 June 2018"
output: html_document
---

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(Hmisc)
library(corrplot)
library(ggplot2)
library(knitr)
library(tidyverse)
library(rmarkdown)
library(ggpubr)
library(ggmap)
library(dplyr)
library(ggpubr)
library(ggmap)
library(ggrepel)
library(plot3D)
```

A rough start

## Introduction:

“Aquatic macrophytes” or “hydrophytes” are terms employed to denote a diverse assemblage of photosynthetic aquatic biota. Typically speaking, these organisms are large enough such that they may be observed unaided, by the naked eye. This group of organisms includes macroalgal members of the divisions Rhodophyta (red algae), Cyanobacteria (blue-green algae), Xanthophyta (yellow-green algae, Chlorophyta (green algae), the Bryophyta (including the liverworts and mosses), the Pteridophyta (including the ferns) as well as the Spermatophyta (commonly known as the seed-bearing plants) (Denny, 1985, Pieterse, 1990). The group is further characterized by vegetative anatomical structures which are confined either perpetually or periodically (at the very least several weeks per annum) to either the upper surface of water bodies, extending through the water surface or alternatively submerged below the surface of the water (Denny, 1985; Pieterse, 1990).
Macrophytes have been identified as key constituents of both wetland and aquatic ecosytems, where they are capable of shaping both the structure and the functioning of aquatic  environments due to the complex roles they exhibit within these habitats (Ervin &Wetzel, 2003; Hrivnak et al. 2009; Tamire and Mengistou 2012) . The macrophytes form the basis of food chains within their environments, while simultaneously contributing to the maintenance and promotion of foodwebs and the provision of vital ecosystem services within aquatic ecosytems (Scheffer and Jeppeson, 2007; Smith, 2011). Furthermore, these organism have been identified as playing a role in the provision of microhabitats, as their leaves, stems and roots, serve as a substrate for periphyton, collectively provide shelter for a wide range of invertebrates, fish, amphibians and reptiles (Timms and Moss, 1984). Amongst many of the other biotic components of aquatic ecosystems, macrophytes have been implicated as a primary factor in both the development and regulation of oxygen content as well as water quality in naturally occurring freshwater bodies (Moore and Ramamoorthy, 1987; Lychagina et al., 1998). The highly specific ecogeological role purported by this group of organisms comes as a direct result of an exceptionally high assimilation potential. This feature is closely correlated with their role in the production of organic matter, which inevitably precipitates and goes on to form bottom sediments. Perhaps another important feature of the macrophytes lies in their ability to accumulate biogenic elements and toxic substances such as heavy metals-which is directly correlated with the abundance of these organisms within any particular aquatic environment (Dikieva and Petrova, 1983; Burdin and Zolotukhina, 1998; Lychagina et al., 1998). The aforementioned properties and capabilities has largely resulted in an increased interest of macrophytes as objects of pollution and biogeochemical indicators within aquatic environments (Boháč, 1987; Lychagina et al., 1998, Osmolovskaya and Kurilenko, 2001; Kurilenko et al., 2002, Kurilenko and Osmolovskaya, 2006), and more importantly their ability to serve as important bioindicators of long term environmental fluctuations in water quality as well as environmental conditions make them indispensable in ecosystem management of freshwater habitats  Lacoul and Freedman, 2006; Solimini et al. 2006).
It is realized that the manner in which macrophytes function within their respective ecosytems is directly correlated with the structural properties and attributes of the population itself, and may be considered in terms of their distribution, species composition, diversity as well as the abundance, all of which are governed by environmental factors such as water temperature, light, pH, depth, nutrients and gas composition, to name but a few (Kual et al, 1978, Pandit, 1984; Barko and Smart, 1986, Cronk and Fennessey, 2001).

The genus Isolepis, is a group of submerged macrophytes which encompasses a total of 76 species, that are typically distributed across cooler tropical and temperate localities and enjoy extensive distributions across Australasia and Africa (Muasya, 1998). Within South Africa, the genus encompasses a total of 30 species, 20 of which have been identified as being endemic to the region (Archer, 2000). The high levels of diversity and endemism exhibited by its members may be attributed to speciation events which may have been catalyzed by climatic changes affecting the entire Cape floristic Region during the Tertiary through Pleistocene period (Linder, 2005). Within South Africa, Isolepis exists as individuals exhibiting an annual growth cycle, typically occurring in ephemeral wetland areas (obligate aquatic organisms), or as perennial species which typically occupy the edges of wetland areas (facultative aquatics), alternatively they  may establish themselves on the surface of shallow freshwater systems (Muasya and Simpson, 2002). The species Isolepsis digitata, is endemic to the Cape Floristic Region, which extends throughout the southwestern regions of the Western Cape, northward towards the Nama-Karoo region, and is characterized by a winter rainfall climate (Archer, 2000). Aside from the morphological characteristics of this species, the available body of information regarding its growth and ecological role within freshwater ecosystems remains scant. Within the Cape Floristic Region, we find that it enjoys distributions between Clanwilliam and Riversdale, within the Western Cape Province, and is commonly found in fast flowing streams, where it forms dense clusters on the surfaces of rocks and flowers during the spring and summer months (Archer, 2000)
Here we investigate the role and autecology of the macrophyte Isolepis digitata, throughout the Krom River ecosystem. The study aims to provide a better understanding regarding the effects of physical environmental parameters on the developmental biology of Isolepis digitata, with particular emphasis being placed on the impact that water parameters such as pH, salinity, depth and gas composition have on biomass accumulation.


Archer, C [in:] Goldblatt, P., & Manning, J. (2000). Cape plants: a conspectus of the Cape flora of South Africa. National Botanical Institute.
Barko, J. W., & Smart, R. M. (1986). Sediment‐related mechanisms of growth limitation in submersed macrophytes. Ecology, 67(5), 1328-1340.
Boháč, J. (1987). AM Nikanorov, AV Zhulidov and AD Pokarzhevskii: Biomonitoring tyazhelykh metallov v presnovodnykh ekosistemakh (Biomonitoring of heavy metals in fresh water ecosystems).–144 pp. Leningrad: Gidrometeoizdat, 1985 (in Russian).

Burdin, K. S., & Zolotukhina, E. Y. (1998). Tyazhelye metally v vodnykh rasteniyakh (akkumulyatsiya i toksichnost’).Denny, P., 1985. The Ecology and Management of African Wetland Vegetation. W. Junk, The Hague
Cronk, J., & Fennessey, M. (2001). Wetland Plants: Biology and Ecology (p. 462). Boca Raton: Lewis Publishers.
Dikieva, D., & Petrova, I. A. Chemical Composition of Macrophytes and Factors Determining Concentrations of Mineral Substances in Higher Aquatic Plants. Gidrobiologicheskie protsessy v vodoemakh, 107-211.
Ervin, G. N., & Wetzel, R. G. (2003). An ecological perspective of allelochemical interference in land–water interface communities. Plant and Soil, 256(1), 13-28.
Hrivnák, R., Oťaheľová, H., & Valachovič, M. (2009). Macrophyte distribution and ecological status of the Turiec River (Slovakia): Changes after seven years. Archives of Biological Sciences, 61(2), 297-306.
Kaul, S. (1982). Community architecture biomass and production in some typical wetlands of Kashmir. Indian journal of ecology.
Kurilenko, V. V., & Osmolovskaya, N. G. (2006). Ecological-biogeochemical role of macrophytes in aquatic ecosystems of urbanized territories (an example of small water bodies of St. Petersburg). Russian Journal of Ecology, 37(3), 147-151.
Lacoul, P., & Freedman, B. (2006). Environmental influences on aquatic plants in freshwater ecosystems. Environmental Reviews, 14(2), 89-136.
Linder, H. P. (2005). Evolution of diversity: the Cape flora. Trends in plant science, 10(11), 536-541.
Lychagina, N. Y., Kasimov, N. S., Lychagin, M. Y., & Kasimov, N. S. (1998). Biogeochemistry of macrophytes of the Volga delta. Geoecology of the Caspian Sea, 4.Kurilenko, V. V., & Osmolovskaya, N. G. (2006). Ecological-biogeochemical role of macrophytes in aquatic ecosystems of urbanized territories (an example of small water bodies of St. Petersburg). Russian Journal of Ecology, 37(3), 147-151.
Moore, J. W., & Ramamoorthy, S. (2012). Heavy metals in natural waters: applied monitoring and impact assessment. Springer Science & Business Media.
Muasya, A. M. (1998). Systematics of the genus Isolepis R. Br.(Cyperaceae). Reading: Muthama Muasya x, 185p.-illus.En Icones, Anatomy and morphology, Keys.Systematics: ANGIOSPERMAE (CYPERACEAE: ISOLEPIS)
Osmolovskaya, N. G., & Kurilenko, V. V. (2001). Biogeochemical aspects of heavy metals phytoindication in urban aquatic ecosystems. Biogeochemical Processes and Cycling of Elements in the Environment, 217-218.
Pandit, A. K. (1984). Role of macrophytes in aquatic ecosystems and management of fresh water resources. J. Environ. Manage.;(United States), 18(1).Pieterse A. H., 1990. Introduction (Chapter 1). In Pieterse A.H., & K. J. Murphy (eds), Aquatic Weeds. Oxford University Press, Oxford, UK: 3–16.

Salt, D. E., Smith, R. D., & Raskin, I. (1998). Phytoremediation. Annual review of plant biology, 49(1), 643-668.
Scheffer, M., & Jeppesen, E. (2007). Regime shifts in shallow lakes. Ecosystems, 10(1), 1-3.
Smith J.E. (2011), Algae [in:] Simberloff, D., & Rejmánek, M. (Eds.). (2011). Encyclopedia of biological invasions (No. 3). University of California Press.
Solimini, A. G., Cardoso, A. C., & Heiskanen, A. S. (2006). Indicators and methods for the ecological status assessment under the Water Framework Directive. Linkages between chemical and biological quality of surface waters Joint Research Centre, European Commission, 1-262.
Tamire, G., & Mengistou, S. (2013). Macrophyte species composition, distribution and diversity in relation to some physicochemical factors in the littoral zone of Lake Ziway, Ethiopia. African Journal of Ecology, 51(1), 66-77.
Timms RM, Moss B. 1984. Prevention of growth of potentially dense phytoplankton populations by zooplankton grazing, in the presence of zooplanktivorous fish, in a shallow wetland ecosystem. Limnol Oceanogr 29: 472-486.


## Aim
To understand the role and autecology of Isolepis digitata along the Krom River.  The objectives are as follows:
To determine the effect of physical parameters on the morphology of Isolepis digitata. This will be done by assessing whether temperature, salinity and pH has an influence on the amount of rhizomes present or if these variables influence the biomass. To assess whether pH varies with depth.

## Study site
Samples were collected along the Krom River at the Kromrivier Cederberg Park (-32.54° S, 19.28° E) in February 2018 between 11am - 2pm. Physical parameters were measured for 2 sites along the Krom river between the top of the river at Disa pool and the middle of the river along a campsite.

## Methods 

### Field 

To determine the physical parameters influencing the distribution and abundance of Isolepis digitata, physical parameters including temperature (°C), pH, electrical conductivity (µS/cm), total dissolved solids (mg/l), salinity (PSU), dissolved O2 (mg/l) and O2 Saturation (%) was measured using a multi-parameter Aquaprobe at each point along the river where an Isolepis digitata sample was removed at each of the two sites. The Aquaprobe was placed at a depth of 3cm below surface waters and remained in the water for approximatly 3minutes until each of the visible recorded values stablised.
A total of 10 samples were collected at each of the two sites, 10 from deep (21-36 cm) top section and 10 from shallow (10-20 cm) middle site water depths. Each Isolepis digitata sample that were present was removed from the bedrock using a 15cm scraper and was collected in a 300 x 300 mm SASS net having a mesh size of 1 mm. The contents of the SASS net were emptied into a plastic tray filled with 1/3 river water, making sure to remove any additional aquatic invertebrates still on the net into the tray. The aquatic invertebrates present on each Isolepis digitata sample was removed into the surrounding water in the plastic tray. Each Isolepis digitata sample was then removed from the plastic tray, placed in a ziplock bag filled with 1/3 river water and stored in the freezer until used in the laboratory. The remaining contents in the plastic tray were placed through a sieve in order to collect the aquatic invertebrates. The collected aquatic invertebrates were washed through a funnel into plastic containers with ethanol, making sure to fill the plastic containers with 1/3 ethanol. At the top site, 10 bedrock samples were also collected to compare the abundance of aquatic invertebrates present on the surrounding bedrock to the abundance present on Isolepis digitata.  
The aquatic invertebrates corresponding to each Isolepis digitata sample were further separated into individual eppendorf tubes filled with ethanol to be analysed at a later stage. 

### Laboratory 

After sample collection, the frozen Isolepis digitata samples were defrosted and laboratory experiments were conducted. The number of rhizomes, the culm length from the top of the soil to the tip of the plant for five culms and the number of spikelets were measured for each Isolepis digitata sample that was collected. All samples were placed in individual brown paper bags and were dried in the oven for a period of 4 days at 40°C. To determine the response of biomass to different environmental conditions, the total mass of each Isolepis digitata sample was recorded after the drying period and the below water mass was separated from the above water mass and was further weighed separately for each Isolepis digitata sample. 


### Data Analysis
After collection, the experimental data was recorded onto a Microsoft© Excel spreadsheet.  The spreadsheet was then converted into comma separated file (CSV) format. The statistical computing software, R Version 3.5.0, was then used to analyse the data and create graphic representations of the information that was collected. T-tests were used in order to determine if significant differences were present between the variables within the dataset. A t-test was used to determine whether there was a significant difference between the macrophyte biomass at a particular depth . A second t-test was used to determine if a significant difference exist between macrophyte biomass and TDS values. In the statistical analyses, values less than that of 0.05, were considered statistically significant.


Themes for graphs
```{r}

theme_AJMS <- function(AfricanJournalOfMarineScience) {
  theme(
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(size = 0.4, fill = NA),
    text = element_text(family = "Arial"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title = element_text(size = 9),
    axis.ticks = element_line(size = 0.5),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(family = "Arial", size = 9),
    legend.text = element_text(family = "Arial", size = 9),
    legend.key = element_blank(),
    legend.key.height = unit(.3, "cm"),
    legend.box.spacing = unit(0.4, "cm"),
    legend.background = element_blank(),
    plot.title = element_text(size = 9, hjust = 0),
    strip.background = element_blank(),
    strip.text = element_text(size = 9),
    axis.ticks.length = unit(-0.2, "cm"),
    axis.text.x = element_text(margin = unit(c(0.25, 0.5, 0.5, 0.5), "cm"), vjust = 0.35, angle = 450),
    axis.text.y = element_text(margin = unit(c(0.5, 0.3, 0.5, 0.5), "cm"))
  )
}
```


Loading the data
```{r load_files1, include=TRUE}
R_assignment_AJ_ <- read_csv("R_assignment(AJ).csv")
Map_coordinates <- read_csv("Map_coordinates.csv")
load("~/Documents/R_assignment/south_africa_coast.RData")
load("~/Documents/R_assignment/sa_provinces.RData")
```


Possible code for mapping of the cederberg area (A.A)
```{r map_func, include=TRUE}
mapping_plot <- function(df){
  mapping <- df %>% 
ggplot(aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  coord_cartesian(xlim = c(12, 36), ylim = c(-38, -26)) +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = df)  +
  geom_label_repel(data =df, aes(x = lon, y = lat, label = Site), 
                   size = 3, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, segment.alpha = 0.4) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
  return(mapping)
}

map_1 <- mapping_plot(df = Map_coordinates)
map_1
```

Looking at biomass and site

```{r}
dat1 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.pH = mean(pH),
            sd.pH = sd(pH))

plot_A <- ggplot(dat1, aes(x = Site, y = mn.pH)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.pH - sd.pH, 
                    ymax = mn.pH + sd.pH)) +
  labs(x = "Site", y = "Average pH") +
  theme_AJMS()

dat2 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Salinity = mean(Salinity),
            sd.Salinity = sd(Salinity))

plot_B <- ggplot(dat2, aes(x = Site, y = mn.Salinity)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Salinity - sd.Salinity, 
                    ymax = mn.Salinity + sd.Salinity)) +
  labs(x = "Site", y = "Average Salinity") +
  theme_AJMS()

dat3 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Temperature = mean(Temperature),
            sd.Temperature = sd(Temperature))

plot_C <- ggplot(dat3, aes(x = Site, y = mn.Temperature)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Temperature- sd.Temperature, 
                    ymax = mn.Temperature + sd.Temperature)) +
  labs(x = "Site", y = "Average Temperature") +
  theme_AJMS()

dat4 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Macbio = mean(Macrophyte_Biomass),
            sd.Macbio = sd(Macrophyte_Biomass))

plot_D <- ggplot(dat4, aes(x = Site, y = mn.Macbio)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Macbio - sd.Macbio, 
                    ymax = mn.Macbio + sd.Macbio)) +
  labs(x = "Site", y = "Average Macrophyte") +
  theme_AJMS()


Final <- ggarrange(plot_A, plot_B,plot_C,plot_D,
                   ncol = 4, nrow = 1,
                   common.legend = TRUE)
Final
```



```{r}
# Comparing macrophyte_biomass and Depth
summary(R_assignment_AJ_$Macrophyte_Biomass)
data <- R_assignment_AJ_
data %>%
  summarise(min_t = min(Macrophyte_Biomass),
            quart_1 = quantile(Macrophyte_Biomass, 0.25),
            med_t = median(Macrophyte_Biomass),
            mean_t = mean(Macrophyte_Biomass),
            quart_3 = quantile(Macrophyte_Biomass, 0.75),
            max_t = max(Macrophyte_Biomass))

A <- ggplot(data, aes(x = Depth, y = Macrophyte_Biomass)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_AJMS()

# Comparing macrophyte biomass with tds
B <- ggplot(data, aes(x = TDS, y = Macrophyte_Biomass)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_AJMS()

combined <- ggarrange(A, B,
                   ncol = 2, nrow = 1,
                   common.legend = TRUE)
combined
```

```{r}
# Areas and adding trend
# Macrophyte Biomass
PLOT1 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = Depth)) +
  geom_point(aes(colour = Site)) +
  geom_smooth(aes(colour = Site), method = "lm") +
  coord_equal(xlim = c(0, 60), ylim = c(0,60)) +
    theme_AJMS()
# Gray shade is the standard error
# angle of line shows the relationship


# Macrophyte biomass and TDS
PLOT2 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = TDS)) +
  geom_point(aes(colour = Site)) +
  geom_smooth(aes(colour = Site), method = "lm") +
  coord_equal(xlim = c(0, 60), ylim = c(0,60)) +
    theme_AJMS()

combined1 <- ggarrange(PLOT1, PLOT2,
                   ncol = 2, nrow = 1,
                   common.legend = TRUE)
combined1
```





