---
title: "Investigating the impact of environmental variables on the biology of *Isolepis Digitata* in the Krom River (Cederberg Park, Western Cape, South Africa"
author: "Amieroh Abrahams, Kim Scholtz and Yameen Badrodien"
date: "20 June 2018"
output: html_document
  word_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: default
    number_sections: yes
    theme: paper
    toc: yes
    toc_float: yes
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: zenburn
    latex_engine: xelatex
mainfont: PT Serif
monofont: PT Mono
language: Australian
sansfont: PT Sans
fontsize: 12pt
---

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(Hmisc)
library(corrplot)
library(ggplot2)
library(knitr)
library(tidyverse)
library(rmarkdown)
library(ggpubr)
library(ggmap)
library(dplyr)
library(corrplot)
library(ggpubr)
library(ggmap)
library(ggrepel)
library(plot3D)
library(Hmisc)
```

## Abstract

*Isolepis digitata* and water specimens were sampled from two sites within the Krom River freshwater ecosystem in the Krom River Cederberg Park, located in the Western Cape Province of South Africa during February 2018. A comparative analysis of the physical (temperature, depth), chemical (sediment, pH, total dissolved oxygen) properties of river water, as well as biological parameters (biomass accumulation of *I.digitata*) revealed the relationships between environmental variable and the biology of *I.digitata*. A comparison of the upper and intermediate sites, in terms of the physico-chemical parameters indicated that sites differed in terms of physical attributes such as temperature, as well as chemical attributes such as pH and total dissolved solids(TDS), while features such as depth and total dissolved oxygen appeared to exhibit no significant differences. Investigation of the biological attributes of *I.digitata*, indicate that the accumulation of biomass exhibited by individuals sampled along the Krom River, were consistent, and exhibited little to no variation between the upper and intermediate sites. Correlation tests between physico-chemical attributes of river water on the biology of *I.digitata*, indicated that biomass accumulation of *I.digitata* within the Krom River, operated independently of measured physical (temperature, depth) and chemical (sediment, pH, total dissolved oxygen) environmental factors, and may possibly be governed by other environmental variables that were not investigated here.

## Introduction 

Freshwater is regarded as perhaps the single most important natural resource required by human beings. The role of freshwater is amplified in impoverished countries, as freshwater ecosystems provide an array of indispensable natural resources which inevitably influence human livelihood and health (Covich et al. 2004). These freshwater ecosystems also support a cornucopia of plant and animal assemblages, contributing significantly to biodiversity, which further contributes to the value of the types of ecosystems in terms of resource provision (Dudgeon et al. 2006; Vié 2008). We find that the proportion of biodiversity sustained by these ecosystems is significantly high, despite the fact that freshwater ecosystems encompass a mere 0.8% of the Earth’s surface area, and constitute 0.01% of the Earth’s freshwater (Gleick et al. 1996). Moreover, it is established that these freshwater ecosystems, serve as habitats to at least 126 000 described species of plants and animals, accounting for at least 7% of all described species of living organisms, which is considerably high given the relatively small area of this type of habitat (Dudgeon et al. 2006; Bailan et al. 2010). Despite the ecological value of these habitats, these types of ecosystems have in recent times, exhibited accelerated rates of biodiversity loss (Dudgeon et al. 2006). 

Based on studies by Dudgeon et al (2006), it was realized that these freshwater ecosystems are currently being subjected to an array of threats which include habitat transformation and degradation, increased levels of water pollution, the overexploitation of biota, as well as a number of impacts that are closely associated with water abstraction and subsequent flow modification coming as a direct consequence of man-made water retention reservoirs. It becomes woefully apparent that factors such as precipitation, temperature and run off patterns, a product of global climate change, may inevitably exacerbate the aforementioned threats and consequently lead to the physical isolation and fragmentation of these types of ecosystems (Dudgeon et al. 2006). These threats coupled with the aforementioned factors have ultimately led to significant decreases in biodiversity within these freshwater ecosystems. According to Darwall et al (2011), at least 20% of amphibian biodiversity, 27% of mollusk diversity and 30 percent of freshwater fish diversity are currently classified as threatened in accordance with the IUCN Red List, which goes on to describe freshwater ecosystems as perhaps the most endangered ecosystems in the globe. Studies by Vié (2008), reveal that freshwater ecosystems serve as habitats to at least 65 000 species of invertebrates, 12 000 species of freshwater fish and 500 species of amphibians. While aquatic plant species have been identified as contributing significantly less to biodiversity within these types of habitats, their presence in freshwater ecosystems have been highlighted given their involvement in the provision of a multitude of ecosystem services as well as playing an integral role in crucial ecological processes (Lacoul & Freedman 2006; Wersal et al. 2006; Bornette & Puijalon 2009).

Studies by Balian et al (2010) have shown that macrophytes encompass a total of 2614 species within freshwater ecosystems and exhibiting low levels of diversity given the fact that they comprise a mere 2% of all described freshwater biota, and represent 1% of all described vascular plant biota. Macrophytes typically enjoy cosmopolitan distributions, with the highest levels of endemism and species diversity being exhibited in temperate regions across the globe (Bailan et al. 2010). It should however be noted that certain tropical and subtropical regions may exhibit significantly high levels of endemism and species richness- these include areas within South America, where at least 81 % of these organism are considered to be endemic, as well as areas within Southern Africa, where they exhibit roughly 20% endemism (Bailan et al. 2010). Within South Africa, we find high diversity of wetland macrophytes, with centers for diversity and endemism being concentrated to areas of the Western Cape and Kwa-Zulu Natal (Darwall et al. 2009). Despite the high levels of diversity within these regions, it is realized that the existing body of information currently available for submerged species of macrophytes remains scant, as the focus of previous studies have generally directed their attention towards floating and emergent species. 

The primary characteristic separating submerged macrophytes from their emergent and floating counterparts lies in the fact that they exist entirely beneath the surface of the water. Submerged macrophytes are characterized by their ability to alter and maintain various components of the freshwater habitat, including the regulation of turbidity within water columns, physico-chemical properties of water while simultaneously acting as carbon reservoirs (Scheffer & Jeppeson 2007; Kennedy et al. 2015). It is realized that the manner in which macrophytes function within their respective ecosystems, may ultimately be governed by the structural attributes of each individual population, as well as factors such as distribution, species composition and diversity (Kaul et al. 1978). Moreover, we find that these aforementioned factors are ultimately established by environmental factors such as the temperature, pH, light, depth, nutrient and gas composition of the waters in which they are distributed (Barko & Smart 1986).  Given the close biophysical relationship between these organisms and their environment, we find that they are often valuable indicators of ecosystem function and health, as macrophyte composition and biomass are directly correlated with certain environmental factors (Lacoul & Freedman 2006)

*Isolepis digitata* (*Cyperaceae*), is water sedge endemic to riverine systems within the Western and Eastern Cape provinces of South Africa (van Ginkel et al. 2011). The species exhibits a perennial growth habit, and is commonly found in dense tufted mats, often attached to  the surfaces of rocks within freshwater ecosystems, and typically found in the higher, fast flowing regions of rivers (Low 2009). The biomass accumulation and growth of this species has been closely linked to temperature, with temperatures above 13°C, promoting both the accumulation of biomass and the rate of growth of the species (Tshilla 2014). *I.digitata* has been shown to play an integral role in the maintenance and regulation of several aspects of freshwater ecosystems - this includes the maintenance of shore integrity (erosion control) while simultaneously maintain the quality of waters in which they are distributed (Tshilla 2014). Despite the significant role it plays in freshwater ecosystems, as well as its endemic status, the species has received very little attention, and as a consequence, we find that the available body of information regarding the autecology of the species remains scant (Low 2009). Studies by Nel et al. (2007) have highlighted the importance of prioritizing the conservation of riverine systems within the Western Cape Province, given the fact that up to 64% of the river ecosystems within the province are currently classified as endangered. The Krom River itself has undergone extensive transformation - a direct consequence of anthropogenic activities that have manifested in the form of altered community structures, decreased biodiversity within freshwater habitats and increased encroachment of invasive riparian vegetation. The aforementioned alterations to these habitats have inevitably led to a decrease in both macrophyte diversity and abundance (Wersal 2006). 

Here we investigate the utility of the submerged macrophyte *I.digitata* and its biology as an indicator of ecosystem health, while highlighting the role that physical and chemical parameters of river water have on its biology. The study seeks to quantify and compare the physical and chemical parameters between upper and intermediate sites within the Krom River; considering factors such as pH, depth (cm), temperature (°C), total dissolved oxygen (mg/l) and total dissolved solids (mg/l). In addition to this, the aforementioned physical and chemical parameters of river water will be correlated with biomass accumulation, in an effort to establish the influence that these environmental factors have on the biology on *I.digitata* individuals within the Krom River system. 

Hypothesis

H0: The accumulation of biomass of *I.digitata* is independant of physical (temeprature and depth) and chemical(pH, total dissolved solid[TDS] and total dissolved oxygen) parameters of the Krom River.

H1: The accumulation of biomass of *I.digitata* is dependant on physical (temperature and depth) and chemical(pH, total dissolved solid[TDS] and total dissolved oxygen) parameters of the Krom River.

## Materials and methods

### Study area
The study was conducted at the Krom River (Figure 1), which is an eastward flowing tributary originating from the Matjies River, within the Olifants-Doring River catchment area and exhibits a total length of approximately 15 km (Marr et al. 2012; Shelton et al. 2017). The Krom River is situated within the Cedarberg, located approximately 280 km outside of Cape, within the Western Cape Province, South Africa (de Moor & Day 2013). The area is marked by the presence of the Cape Fold Mountains, comprising shale and quartzitic sandstone which results in rivers of this region exhibiting waters which are typically oligotrophic in nature, and slightly acidic in terms of its pH (Goldblatt & Manning 2002; de Moor & Day 2013). The region is further characterized by a Mediterranean climate and thus, the region typically experiences hot dry summer months, followed by winter months which are cold and wet (Shelton et al. 2017). As such, the vast majority of rainfall experienced within this region occurs during the winter months, with the average maximum rainfall falling in the range of 250mm per annum (Garrow & Marr, 2012). This study identified two sites within the study area, that being, the upper site and the intermediate site. The sites were separated on the basis of depth, with the upper site exhibiting a greater depth when compared to the intermediate site, which was considerably shallower. The upper site was located above the Disa Pool, while the intermediate site was located directly adjacent to the campsite within the Krom River Cedarberg Park. All samples collected from the study area were recorded using a hand held GPS (Garmin eTrex 10). All sampling took place during February 2018, over a 4 day period, between 11h00 and 14h00.

### Species distribution and aquatic habitat characterization

A wide array of variables, relating to the physico-chemical properties of river water was recorded at each of the two sites. These variables included the temperature (°C), pH, electrical conductivity (μS cm-1), total dissolved solids (TDS: ppm) and salinity (PSU) were quantified in situ using a Hanna® HI 98129 pocket multi-function meter (HANNA Instruments, Inc.). A total of 10 replicates were taken at both the upper and intermediate sites, at a depth of 3 cm below the water surface, for a total of 3 minutes, or until the measurement appearing on the apparatus stabilized.   

### Biomass quantification

The biomass of *I. digitata* was established through the collection of specimens from both the upper and the intermediate sites within the study area. A total of 10 specimens were collected from either of the two sites, from a 25 cm x 25 cm quadrat. Following collection, specimens were identified to species level using a field guide. Collected submerged macrophytes were placed in zip-lock bags filled with 1/3 river water, and frozen at -12°C for further analyses in the laboratory. In the lab, macrophytes were thawed at room temperature, and rinsed using a sieve in an effort to remove any ambient material or sediment which may have been present on the specimens. The macrophytes were then dehydrated at 40°C for a period of 72 hours, using the Memmert 854 Schwabach® Model UL80 oven. The biomass (g) of dehydrated specimens was then measured to two decimal places using the Mettler Toledo® New Classic SG Model ML802E scale.  

### Data modification

Once measurements had been established for macrophyte and water specimens, the data was extrapolated (within the given range of the datasets for both upper and intermediate sites). This yielded a total of 40 pseudoreplicates per site, raising the total number of replicates for either site to n = 50. Pseudoreplication was performed in an effort to widen the range and variation of the values obtained from physical samples, thereby facilitating the statistical analysis of the data. 

### Statistical analysis

All statistical analyses conducted, were performed using R. v. 3.4.2. In all of the tests conducted, results were considered significant at p< 0.05. Prior to statistical analyses, the *Shapiro-Wilks Test* was employed to establish the normality of the data. The results of the *Shapiro-Wilks Test* were further corroborated by assessing the homoscedasticity of the data. Homoscedasticity of the data was established through the comparative analyses of variances of each of the variables. Where these differences were less than four or more times in size, it was assumed that the data was normally distributed. Where differences were greater than four or more times, we were able to conclude that the data was non-normally distributed. Analysis of the data using the aforementioned assessments revealed that the data collected relating to both the biomass of *I.digitata*, as well as the physico-chemical properties of river water, were in fact non-normally distributed. Given the nature of the variables being employed, differences between these variables were established using the *Wilcox Rank-Sum Test*. The *Wilcox Rank-Sum Test* is a non-parametric version of the two sample t-test. The *Wilcox Rank-Sum Test* was used to compare physico-chemcial environmental variables, as well as differences in biomass accumulation of collected *I.digitata* specimens between the selected upper and intermediate sites within the Krom River. Boxplots were constructed to provide a graphical representation of distributions relating to both the biological parameters of *I.digitata* (biomass) as well as physico-chemical properties of river water.The relationships between these variables were also graphically represented with a correlation matrix. Non-parametric correlation tests were conducted in an effort to establish the relation and relative impact of each of the aforementioned physico-chemcial impacts had on biological variables (biomass) of *I.digitata*. Given the non-parametric nature of the data collected, the *Spearman Rank Correlation* was employed for the aforementioned effort. *The Spearman Rank Correlation* is the non-parametric version of the Pearsons product-moment correlation.

Creating a theme for all the graphs used in this script: This theme was created in order to meet all the South African Journal of Marine Science visualisation requirements.

```{r}

theme_AJMS <- function(AfricanJournalOfMarineScience) {
  theme(
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(size = 0.4, fill = NA),
    text = element_text(family = "Arial"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title = element_text(size = 9),
    axis.ticks = element_line(size = 0.5),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(family = "Arial", size = 9),
    legend.text = element_text(family = "Arial", size = 9),
    legend.key = element_blank(),
    legend.key.height = unit(.3, "cm"),
    legend.box.spacing = unit(0.4, "cm"),
    legend.background = element_blank(),
    plot.title = element_text(size = 9, hjust = 0),
    strip.background = element_blank(),
    strip.text = element_text(size = 9),
    axis.ticks.length = unit(-0.2, "cm"),
    axis.text.x = element_text(margin = unit(c(0.25, 0.5, 0.5, 0.5), "cm"), vjust = 0.35, angle = 450),
    axis.text.y = element_text(margin = unit(c(0.5, 0.3, 0.5, 0.5), "cm"))
  )
}
```

At this point it was required that the dataset for the study was imported into R Studio 1.1.423. This data comprised teh *in situ* samples collected from the study site, along with the generated pseudo-replicates, as well as co-oordinates needed for generating localised maps of the study area. 

```{r load_files1, include=TRUE}
R_assignment_AJ_ <- read_csv("R_assignment(AJ).csv")
Map_coordinates <- read_csv("Map_coordinates.csv")
load("sa_provinces.RData")
load("south_africa_coast.RData")
```

Map showing the Cederberg region located within South Africa:

```{r, include=TRUE, fig.cap = "Figure 1: A satellite map of the study area located within the Krom River, Krom River Cederberg Park, in the Western Cape province of South Africa"}

mapping_plot <- function(df){
  mapping <- df %>% 
ggplot(aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  coord_cartesian(xlim = c(12, 36), ylim = c(-38, -26)) +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = df)  +
  geom_label_repel(data =df, aes(x = lon, y = lat, label = Site), 
                   size = 3, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, segment.alpha = 0.4) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
  return(mapping)
}

map_1 <- mapping_plot(df = Map_coordinates)
map_1

map1<- get_map(location = c(lon = 19.275, lat = -32.535),zoom = 15, maptype = 'satellite')

trial1 <- ggmap(map1) +
  geom_point(data = Map_coordinates, aes(x = lon, y = lat),
             colour = "red", size = 4) +
  labs(x = "Lat", y = "Lon", title = "Map showing the Cederberg region") +
  theme(plot.title = element_text(hjust = 0.5))
 

new1 <- trial1 +
  geom_text(data = Map_coordinates, aes(lon, lat, label = Site),
            hjust = 1.09, vjust = 0.5,
            size = 5, colour = "white")
plot(new1)

```

The next step was to visualise the data in the form of boxplots, which illustrated the physico-chemcial parameters of either site. This gave us an idea of the variation in environmental variables of both the upper and intermediate sites within the study area.

```{r, include=TRUE, fig.cap = "Figure 2: Boxplots illustrating the variation in physical( temperature [Degrees Celsius] and depth[cm]), chemical (pH, total dissolved solids (TDS: mg/l) and total dissolved oxygen (mg/l) between the selected upper and intermediate sites of the Krom River, Krom River Cederberg Park."}
dat1 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.pH = mean(pH),
            sd.pH = sd(pH))

plot_A <- ggplot(dat1, aes(x = Site, y = mn.pH)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.pH - sd.pH, 
                    ymax = mn.pH + sd.pH)) +
  labs(x = "Site", y = "pH") +
  theme_AJMS()

dat2 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Dissolved_oxygen = mean(Dissolved_oxygen),
            sd.Dissolved_oxygen = sd(Dissolved_oxygen))

plot_B <- ggplot(dat2, aes(x = Site, y = mn.Dissolved_oxygen)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Dissolved_oxygen - sd.Dissolved_oxygen, 
                    ymax = mn.Dissolved_oxygen + sd.Dissolved_oxygen)) +
  labs(x = "Site", y = "Dissolved Oxygen (mg/l)") +
  theme_AJMS()

dat3 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Temperature = mean(Temperature),
            sd.Temperature = sd(Temperature))

plot_C <- ggplot(dat3, aes(x = Site, y = mn.Temperature)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Temperature- sd.Temperature, 
                    ymax = mn.Temperature + sd.Temperature)) +
  labs(x = "Site", y = "Temperature (°C)") +
  theme_AJMS()

dat4 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Macbio = mean(Macrophyte_Biomass),
            sd.Macbio = sd(Macrophyte_Biomass))

plot_D <- ggplot(dat4, aes(x = Site, y = mn.Macbio)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Macbio - sd.Macbio, 
                    ymax = mn.Macbio + sd.Macbio)) +
  labs(x = "Site", y = "Macrophyte Biomass (g)") +
  theme_AJMS()

dat5 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Depth = mean(Depth),
            sd.Depth = sd(Depth))

plot_E <- ggplot(dat5, aes(x = Site, y = mn.Depth)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Depth - sd.Depth, 
                    ymax = mn.Depth + sd.Depth)) +
  labs(x = "Site", y = "Depth (cm)") +
  theme_AJMS()

dat6 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.TDS = mean(TDS),
            sd.TDS = sd(TDS))

plot_F <- ggplot(dat6, aes(x = Site, y = mn.TDS)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.TDS - sd.TDS, 
                    ymax = mn.TDS + sd.TDS)) +
  labs(x = "Site", y = "TDS (mg/l)") +
  theme_AJMS()


Final <- ggarrange(plot_A, plot_B,plot_C,plot_D,plot_E, plot_F,
                   ncol = 3, nrow = 2,
                   common.legend = TRUE)
Final
```

The next step was to create a visualisation of the relationship between the biomass of *I.digitata* and each of the physico-chemical paramters of river water for both the upper and intermediate sites.

```{r, include=TRUE, fig.cap = "Figure 3: Scatterplots exhibiting the realtionship between physical (temperature[degrees celcius] and depth [cm] and chemcial(pH, TDS [mg/l] and total dissolved oxygen[mg/l] on the accumulation of biomass in *I.digitata* individuals within the Krom River, Krom River Cederberg Park))"}
# Areas and adding trend
# Macrophyte Biomass
PLOT1 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = Depth)) +
  geom_point() +
  coord_equal(xlim = c(0, 60), ylim = c(0,60)) +
  labs(x = "Depth (cm)", y ="Macrophyte Biomass (g)") +
    theme_AJMS()
PLOT1

# angle of line shows the relationship
# Macrophyte biomass and TDS
PLOT2 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = TDS)) +
  geom_point() +
  coord_equal(xlim = c(0, 60), ylim = c(0, 60)) +
  labs(x = "TDS (mg/l)", y = "Macrophyte Biomass (g)") +
    theme_AJMS()
PLOT2

# angle of line shows the relationship

# Macrophyte biomass and pH
PLOT3 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = pH )) +
  geom_point() +
  coord_equal(xlim = c(0, 10), ylim = c(0,60)) +
  labs(x = "pH", y = "Macrophyte Biomass (g)") +
    theme_AJMS()
PLOT3
# angle of line shows the relationship

# Macrophyte biomass and temperature
PLOT4 <- ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = Temperature)) +
  geom_point() +
  coord_equal(xlim = c(10, 30), ylim = c(0, 60)) +
  labs(x = "Temperature (Deg Celcius)", y ="Macrophyte Biomass (g)" ) +
    theme_AJMS()
PLOT4
# angle of line shows the relationship

# Macrophyte biomass and dissolved oxygen
PLOT5 <-ggplot(data = R_assignment_AJ_, aes(y = Macrophyte_Biomass , x = Dissolved_oxygen)) +
  geom_point() +
  coord_equal(xlim = c(0, 10), ylim = c(0,60)) +
  labs(x ="Dissolved Oxygen (mg/l)" , y ="Macrophyte Biomass (g)" ) +
    theme_AJMS()

PLOT5

combined1 <- ggarrange(PLOT1, PLOT2,PLOT3, PLOT4,PLOT5, 
                   ncol = 2, nrow = 3,
                   common.legend = TRUE)
combined1

```

Calculating the combined average (upper and intermdiate sites) for each of the physico-chemical and biological parameters (Macrophyte biomass).

```{r, include=FALSE}

R_assignment_AJ_ %>%
  group_by(Site) %>% 
summarise(mean_temp = mean(Temperature),
          mean_pH = mean(pH),
          mean_depth = mean(Depth),
          mean_TDS = mean(TDS),
          mean_mac = mean(Macrophyte_Biomass),
          mean_diss = mean(Dissolved_oxygen))
```

Table 1: Summary of averaged physico-chemical and biological parameters for the upper and intermediate sites located along the Krom River.
--------------------------------------------------------------------------------------------------------------------------------------
                Temperature (°C)   Depth (cm)    pH      TDS (mg/l)  Macrophyte Biomass (g)   Dissolved oxygen (mg/l)  Salinity (PSU)
-------------  -----------------  -----------  -------  ------------ ---------------------- -------------------------  -------------  
                

Middle                   20.58         24.28    6.94      32.70             18.80                   6.37                  0.01

Upper                    20.89         24.90    5.69      26.96             17.07                   6.60                  0.02

--------------------------------------------------------------------------------------------------------------------------------------

Now we test to check the normality of our the dataset, which includes *in situ* measurements as well as the pseudoreplicated measurements.

```{r}

R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(norm_temp = as.numeric(shapiro.test(Temperature)[2]),
            norm_TDS = as.numeric(shapiro.test(TDS)[2]),
            norm_pH = as.numeric(shapiro.test(pH)[2]),
            norm_diss = as.numeric(shapiro.test(Dissolved_oxygen)[2]),
            norm_mac = as.numeric(shapiro.test(Macrophyte_Biomass)[2]),
            norm_depth = as.numeric(shapiro.test(Depth)[2]))

# Testing homoscedasity

R_assignment_AJ_ %>%
group_by(Site) %>%
summarise(sample_temp = var(Temperature),
          sample_pH = var(pH),
          sample_TDS = var(TDS),
          sample_mac = var(Macrophyte_Biomass),
          sample_depth = var(Depth),
          sample_diss = var(Dissolved_oxygen))
```

In order to estab lish the distribution of the data, the *Shapiro-Wilks Test* was employed. The test provides outputs of a w and p value. For our interest of establishing the distribution of the the data only the p-value was taken into account. Where the p-value is above 0.05, it is assumed that the data exhibits a normal distribution. Given the non-normal nature of teh data, conventional independant t-tests could not be employed. Instead, a non-parametric version of teh independant t-tesy, namely the *Wilcox Rank-Sum Test*, was used. In order to perform this test in R, we may again use the compare_means() functuon and specify the test which we would like to use. Here we used the the *Wilcox Rank-Sum Test* to assess the differences in physio-chemical and biological parameters between the upper and intermediate sites

```{r, include=FALSE}

compare_means(Macrophyte_Biomass ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
compare_means(Temperature ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
compare_means(pH ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
compare_means(Dissolved_oxygen ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
compare_means(Depth ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
compare_means(TDS ~ Site, data = R_assignment_AJ_, method = "wilcox.test")
```

Table 2: Statistical outputs (p-values) of the *Wilcox Rank-Sum Test* conducted on the upper and intermediate sites in an effort to establish differences in physical (temperature[degrees celcius], depth [cm]), chemical (pH, total dissolved solids [TDS: mg/l] and total dissolved oxygen[mg/l] and biological parameters [biomass: g]).
----------------------------------------------------------------------------------
                                   Group 1   Group 2      p        
-------------  ---------------- ---------- -------  -----------    
                

Temperature (°C)                   Upper     Middle    0.0010   

Depth (cm)                         Upper     Middle    0.4430

pH                                 Upper     Middle     1.31e-17	

TDS (mg/l)                         Upper     Middle     3.87e-14	

Macrophyte Biomass (g)             Upper     Middle     0.0502    

Dissolved oxygen (mg/l)            Upper     Middle     0.0670
-----------------------------------------------------------------------------------

The *Spearman Rank Correlation*, a non-parametric test, was conducted in order to collectively establish the relationship between each of the environmental variables on biomass accumulation within the study site.

```{r, include=FALSE}

cor.test( ~ Temperature + Macrophyte_Biomass,
         data=R_assignment_AJ_,
         method = "spearman")

cor.test( ~ Depth + Macrophyte_Biomass,
         data=R_assignment_AJ_,
         method = "spearman")

cor.test( ~ pH + Macrophyte_Biomass,
         data=R_assignment_AJ_,
         method = "spearman")

cor.test( ~ Dissolved_oxygen + Macrophyte_Biomass,
         data=R_assignment_AJ_,
         method = "spearman")

cor.test( ~ TDS + Macrophyte_Biomass,
         data=R_assignment_AJ_,
         method = "spearman")
```

Table 3: Statistical outputs of the *Spearman Rank Coefficient Correlation Test* depicting the relationship of environmental variables and biomass accumulation in *I.digitata*
------------------------------------------------------------------------------------------------------------------------------------
                       Groups                                 S              p-value          
--------------------------------------------------------  ---------------- --------------- 

Temperature (°C) and Macrophyte Biomass (g)                  179140            0.46     

Depth (cm)  and Macrophyte Biomass (g)                       198680            0.06       

pH and Macrophyte Biomass (g)                                144160            0.18                  

TDS (mg/l) and Macrophyte Biomass (g)                        153950            0.45
  
Dissolved oxygen (mg/l)  and Macrophyte Biomass (g)          182770            0.34
-------------------------------------------------------------------------------------------------------------------------------------

Here we create a correlation matrix as a means of graphically illustrating the relationships between physical and chemical environmental variables on the accumulation of biomass of the submerged macrophyte *I.digitata*

```{r}

corr_data <- read_csv("~/Documents/R_assignment/corr_data.csv")

res2 <- rcorr(as.matrix(corr_data))
res2

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
  )
}

cor3<-rcorr(as.matrix(corr_data, use = "complete.obs"))
flattenCorrMatrix(cor3$r, cor3$P)
```


Plotting the correlation matrix, this provides a visual representation of all the parameters used within this study.

```{r,include=True, fig.cap = "Figure 4: Correlation matrix of environmental variables on the accumulation of biomass of *I.digitata* within the upper and intermediate sites of the Krom River, Cederberg Park)}

try1 <- cor(corr_data, use = "complete.obs")
try1

# Making a visualisation
corrplot(try1)
```
## Discussion

Studies employing a functional method with regards to community ecology typically direct their attention to the relationships exhibited between abiotic factors (physical and chemical parameters) and the composition and richness of macrophyte growth forms within freshwater ecosystems (e.g. Akasaka et al. 2010; Alahuta & Heino 2013). Hudon et al. 2000 suggest that the potential impacts of these environmental variables on the accumulation and variation in biomass amongst macrophytes, as basic as it may be, often goes under looked and is grossly underestimated. The application of these types of relationships have become more relevant in recent years, as the utility and value of these species as indicator species of both ecosystem function and health has become more apparent (Lacoul and Freedman, 2006). The main findings of this study reveal that the chemical and physical variables selected in this study, are incapable of explaining variation in the biomass of the submerged macrophyte *I.digitata* along the two selected sites (upper and intermediate) within the site. Considering these findings, it is realized that our hypothesis that the biology of the submerged macrophyte *I.digitata* was is dependant on physico-chemical attributes of its aquatic environment is subsequently rejected. The approaches employed in this study revealed the following: Firstly, despite significant differences in physico-chemical parameters (temperature, pH and TDS) between the selected upper and intermediate sites, no significant differences in the biomass of *I.digitata* were observed between the two sites (*Wilcox Rank-Sum Test* p > 0.05) (see Fig. 2 and Table 2).Moreover, the results emphasize that the biomass accumulation of the submerged macrophyte *I.digitata*, is weakly associated with physico-chemical parameters investigated in this study (See Fig. 3 and 4). *Spearman Rank Correlation* tests revealed that biomass accumulation of *I.digitata* exhibited weak associations with physical parameters such as depth and temperature, and with chemical parameters such as pH, total dissolved solid (TDS) and total dissolved oxygen (*Spearman Rank Correlation* p > 0.05) (See Table: 3). These findings may be corroborated with mean values obtained for each of the physical and chemical parameters obtained from either site (see Table: 1) 

The findings of this study come in direct contrast with previous literature relating to the effects of environmental variables on the variation of macrophyte life form biomass. Studies by Madsen et al. 2001 emphasize the impact of chemical parameters such as total dissolved solid (TDS) on the accumulation of biomass, as particle resuspension induced by water movement has the capacity to limit the penetration of light through the water column, and may serve as a limiting factor to photosynthetic activity in submerged macrophytes. Furthermore, studies by Junk and Piedade 1993, have demonstrated the effects of increased water depth on both net primary productivity and subsequent biomass accumulation- which they attribute to the horizontal expansion of macrophytes in deeper waters. In terms of pH, we find that studies conducted by Takahashi and Asaeda (2012), reveal that there is an increased rate of photosynthetic activity and subsequent biomass accumulation in waters that are slightly acidic. The rationale behind this statement lies in the fact that acidic water inevitably shifts the equilibrium balance of bicarbonate (HCO¬¬¬¬¬3¬) free CO2, which results in a higher amount of free CO2 ¬being made available for uptake by submerged macrophytes. Moreover, Takahashi and Asaeda (2012) have reported the role of temperature on submerged macrophyte biomass accumulation, which states that the physical attributes such as temperature have the capacity to affect chemical parameters such as dissolved oxygen, where higher water temperatures may lead to a decreased ability of gases to remain in solution (Takahashi & Asaeda 2012) 

Despite little to no influence of physico-chemical parameters on the accumulation of *I.digitata* biomass, our results indicate that these parameters differed between the upper and intermediate sites along the Krom River.Significant differences were observed for physical parameters such as temperature (*Wilcox Rank-Sum Test* p = 0.001), as well as chemical parameters such as pH and total dissolved solids (TDS) (*Wilcox Rank-Sum Test* p < 0.001). Physical (depth) and chemical (dissolved oxygen) appear to be more or less consistent between the two sites (Fig. 2). Statistical analyses on the aformentioned physical and chemical parameters revealed that no significant differences existed between the two sites (*Wilcox Rank-Sum Test* p > 0.05).

The indeterminacy of this study, regarding the influence of environmental variables on the biology of *I.digitata*, may be attributed to the employment of pseudoreplicates in the collected dataset. In addition to this results were constrained by limiting the sampling sites to two locations. This ultimately resulted in little to no variation in the data employed for statistical analysis. As a result of this we find that the results of our analysis are inadequate in explaining the relationship that both physical and chemical environmental parameters have on the accumulation of biomass in *I.digitata*. Despite a weakened potential of the data as a result of pseudoreplication, data was still capable of revealing differences in the chemical and physical parameters of river water between the two sites. Future studies would need to include a greater number of sampling sites in conjunction with a higher number of replicates derived from each of the sites, if we wish to better understand the dynamics of environmental variables of the biology of the submerged macrophyte *I.digitata*.


## References


Akasaka M, Takamura N, Mitsuhashi H, Kadono Y. 2010. Effects of land use on aquatic macrophyte diversity and water quality of ponds. *Freshwater Biology* 55(4): 909–922.

Alahuhta J, Heino J. 2013. Spatial extent, regional specificity and metacommunity structuring in lake macrophytes. *Journal of Biogeography* 40(8): 1572–1582.

Balian E, Harrison I, Butchart S, Chambers P, Cordeiro J, Cumberlidge N, de Moor F, Gascon C, James H, Kalkman V, van Dijk PP, Yeo D. 2010. Chapter 2. Diversity of Species in Freshwater Systems. In: RA Mittermeier, T Farrell, IJ Harrison, AJ Upgren & T Brooks (eds), Freshwater: the essence of Life. China: *Earth in Focus Editions*. pp 50–89.

Barko JW, Smart RM. 1986. Sediment‐related mechanisms of growth limitation in submersed macrophytes. *Ecology* 67(5): 1328–1340.

Bornette G, Puijalon S. 2009. Macrophytes: Ecology of Aquatic Plants. Chichester: John Wiley & Sons, Ltd.

Covich AP, Austen MC, Bärlocher F, Chauvet E, Cardinale BJ, Biles CL, Inchausti P, Dangles O, Solan M, Gessner MO, Statzner B. 2004. The role of biodiversity in the functioning of freshwater and marine benthic ecosystems. *BioScience* 54: 767–775.

Darwall WR, Holland RA, Smith KG, Allen D, Brooks EG, Katarya V, Pollock CM, Shi Y, Clausnitzer V, Cumberlidge N, Cuttelod A. 2011. Implications of bias in conservation research and investment for freshwater species. *Conservation Letters* 6: 474–482.

de Moor FC, Day J. 2013. Aquatic biodiversity in the mediterranean region of South Africa. *Hydrobiologia* 719: 237–268.

Dudgeon D, Arthington AH, Gessner MO, Kawabata ZI, Knowler DJ, Lévêque C, Naiman RJ, Prieur-Richard AH, Soto D, Stiassny ML Sullivan CA. 2006. Freshwater biodiversity: importance, threats, status and conservation challenges. *Biological review* 81: 163–182.

Garrow C, Marr S. 2012. Swimming on the edge of extinction. The perilous state of the indigenous freshwater.

Gleick PH. 1996. Basic water requirements for human activities: meeting basic needs. *Water international* 21: 83–92.

Goldblatt P, Manning JC. 2002. Plant diversity of the Cape region of southern Africa. *Annals of the Missouri Botanical Garden* 89: 281–302.

Hudon C, Lalonde S, Gagnon P. 2000. Ranking the effects of site exposure, plant growth form, water depth, and transparency on aquatic plant biomass. *Canadian Journal of Fisheries and Aquatic Sciences* 57(S1): 31–42.

Junk WJ, Piedade MTF. 1993. Biomass and primary-production of herbaceous plant communities in the Amazon floodplain. *Hydrobiologia*, 263(3):155–162. 

Kaul S. 1982. Community architecture biomass and production in some typical wetlands of Kashmir. *Indian journal of ecology* 54: 767–775.

Kennedy MP, Lang P, Grimaldo JT, Martins SV, Bruce A, Hastie A, Lowe S, Ali MM, Sichingabula H, Dallas H, Briggs J. 2015. Environmental drivers of aquatic macrophyte communities in southern tropical African rivers: Zambia as a case study. *Aquatic Botany* 124: 19–28.

Lacoul P, Freedman B. 2006. Environmental influences on aquatic plants in freshwater ecosystems. *Environmental Reviews* 14: 89–136.

Low B. 2009. Rivers running through-examining the botanical hearts of ten Cape rivers. *Veld & Flora* 95: 198–201.

Madsen JD, Chambers PA, James WF, Koch, Westlake EW. 2001. The interaction between water movement, sediment dynamics and submersed macrophytes. *Hydrobiologia* 444(2): 71–84.

Marr SM, Impson ND, Tweddle D. 2012. An assessment of a proposal to eradicate non-native fish from priority rivers in the Cape Floristic Region, South Africa. *African Journal of Aquatic Science* 37: 131–142.

Nel JL, Roux DJ, Maree G, Kleynhans CJ, Moolman J, Reyers B, Rouget M, Cowling RM. 2007. Rivers in peril inside and outside protected areas: a systematic approach to conservation assessment of river ecosystems. *Diversity and Distributions* 13: 341–352.

Scheffer M, Jeppesen E. 2007. Regime shifts in shallow lakes. *Ecosystems* 10: 1–3.

Shelton J, Weyl O, van Der Walt J, Marr S, Impson D, Maciejewski K, Tye D, Dallas H, Esler K. 2017. Effect of an intensive mechanical removal effort on a population of non‐native rainbow trout Oncorhynchus mykiss in a South African headwater stream. *Aquatic Conservation: Marine and Freshwater Ecosystems* 27: 1015–1055.

Takahashi K, Asaeda T. 2012. Effect of pH decline by acidified spring water on the growth of a submerged macrophyte, Egeria densa (Hydrocharitaceae) in the Kurohashi River, Japan. *Journal of Ecology and The Natural Environment* 4(4): 109–113.

Tshiila AA. 2014. Isolepis. Available at www.pza.sanbi.org/isolepis [accessed 29 August 2017].\

Vié JC. 2008. The IUCN Red list: a key conservation tool. In W: Vié JC, Hilton-Taylor C & SN Stuart (Eds.), The 2008 Review of The IUCN Red list of Threatened Species (pp 1–6). Gland, Switzerland: IUCN.  

van Ginkel CE, Glen RP, Gordon-Gray KD, Cilliers CJ, Muasya M, Van Deventer PP. 2011. Easy identification of some South African wetland plants (grasses, restios, sedges, rushes, bulrushes, eriocaulons and yellow-eyed grasses). Water Research Commission, Pretoria. pp 110.

Wersal RM, Madsen JD, McMillan BR, Gerard PD. 2006. Environmental factors affecting biomass and distribution of Stuckenia pectinata in the Heron Lake System, Minnesota, USA. *Hydrobiologia* 26: 313–321.

## Acknowledgements 

All collaborators have contributed equally to the report. The data repository can be found at https://github.com/AmierohAbrahams/R_assignment


## Appendix I - Figures 

```{r}
plot(new1)
```

```{r}
Final
```

```{r}
combined1
```


# Appendix II - Tables 

```{r}
Table 1
```

```{r}
Table 2
```


```{r}
Table 3
```

