---
title: "Assessing the ecological role of Isolepis digitata Schrad. along the Krom River, Cedarberg, Western Cape"
author: "Amieroh, Kim and Yameen"
date: "14 June 2018"
output:
  word_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: default
    number_sections: yes
    theme: paper
    toc: yes
    toc_float: yes
  pdf_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: zenburn
    latex_engine: xelatex
mainfont: PT Serif
monofont: PT Mono
language: Australian
sansfont: PT Sans
fontsize: 12pt
---

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(Hmisc)
library(corrplot)
library(ggplot2)
library(knitr)
library(tidyverse)
library(rmarkdown)
library(ggpubr)
library(ggmap)
library(dplyr)
library(ggpubr)
library(devtools)
library(ggsn)
library(ggthemes)
library(e1071)
```

A very rough, very basic start.


## Introduction:

Submerged macrophytes are aquatic plants with vegetative structures that are predominantly submerged and are integral to the ecological functioning of aquatic environments. They function as producers, are important sources of food and refuge, provide habitats for aquatic invertebrates, zooplankton and fish. In addition to this they alter water movement, sediment and nutrient dynamics. The distribution of macrophytes is determined by abiotic factors such as depth, nutrients, sediment type, pH and most importantly light. The submerged macrophyte, Isolepis, belongs to the sedge family Cyperaceae and consists of annuals and short-lived perennial plants having a grass-like appearance. Isolepis digitata is a tufted perennial with minute roots, ascending whitish rhizomes and green leaves. The genus Isolepis is well represented in South Africa with Isolepis digitata being endemic to the Cape region having its highest abundance in the winter months. It is found between Clanwilliam and Riversdale in the Western Cape in fast flowing streams, forming dense clusters on rock surfaces and flowers between September and January3. Very little is known about of this species, therefore this study focuses on Isolepis digitata and describes its morphology and the ecological role in relation to the environmental characteristics.

## Aim
To understand the role and autecology of Isolepis digitata along the Krom River.  The objectives are as follows:
To determine the effect of physical parameters on the morphology of Isolepis digitata. This will be done by assessing whether temperature, salinity and pH has an influence on the amount of rhizomes present or if these variables influence the biomass. To assess whether pH varies with depth.

## Methods 

Physical parameters including temperature (°C), pH, electrical conductivity (µS/cm), total dissolved solids (mg/l), salinity (PSU), dissolved O2 (mg/l) and O2 Saturation (%) was measured using a multi-parameter Aquaprobe at each point along the river where Isolepis digitata samples were removed.
The target species was collected from each site at two depths (n=5 for each depth). Depths between 21- 36 cm was classified as deep and depths between 10-20 cm was classified as shallow. The number of rhizomes, the culm length (top of the soil to the tip of the plant) for five individual culms and the number of spikelets were measured for each sample. To determine the response of biomass to different environmental conditions, the total biomass of each sample was recorded with the roots and culms being separated and weighed individually for each sample. 
The amount of organic content stored in each sample was measured by combustion using an ashing oven to reflect the role of this species for carbon storage.  The aquatic invertebrates that were present on each sample was collected in the field and preserved in ethanol to represent the role of this species as a refugia for animals.

## Statistical analyses

A correlation matrix was constructed using R statistical software to correlate each of the biological variables (number of rhizomes, culm length and total biomass) to each of the physical variables (temperature (°C), pH, electrical conductivity (µS/cm), total dissolved solids (mg/l), salinity (PSU), dissolved O2 (mg/l) and O2 Saturation (%)). A single factor ANOVA test was included to compare the biomass of samples at deep and shallow depths between sites....... we need to add more to this (A.A).

## Data analysis

Themes for graphs
```{r}

theme_AJMS <- function(AfricanJournalOfMarineScience) {
  theme(
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(size = 0.4, fill = NA),
    text = element_text(family = "Arial"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.title = element_text(size = 9),
    axis.ticks = element_line(size = 0.5),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(family = "Arial", size = 9),
    legend.text = element_text(family = "Arial", size = 9),
    legend.key = element_blank(),
    legend.key.height = unit(.3, "cm"),
    legend.box.spacing = unit(0.4, "cm"),
    legend.background = element_blank(),
    plot.title = element_text(size = 9, hjust = 0),
    strip.background = element_blank(),
    strip.text = element_text(size = 9),
    axis.ticks.length = unit(-0.2, "cm"),
    axis.text.x = element_text(margin = unit(c(0.25, 0.5, 0.5, 0.5), "cm"), vjust = 0.35, angle = 450),
    axis.text.y = element_text(margin = unit(c(0.5, 0.3, 0.5, 0.5), "cm"))
  )
}
```

Loading the data
```{r load_files1, include=TRUE}
R_assignment_AJ_ <- read_csv("R_assignment(AJ).csv")
```


Possible code for mapping of the cederberg area (A.A)
```{r map_func, include=TRUE}
mapping_plot <- function(df){
  mapping <- df %>% 
ggplot(aes(x = lon, y = lat)) +
  geom_polygon(data = south_africa_coast, aes(group = group), fill = "grey70") +
  coord_cartesian(xlim = c(12, 36), ylim = c(-38, -26)) +
  geom_path(data = sa_provinces, aes(group = group)) +
  geom_point(data = df)  +
  geom_label_repel(data =df, aes(x = lon, y = lat, label = site), 
                   size = 3, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, segment.alpha = 0.4) +
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())
  return(mapping)
}
```

Or we could use google map version to plot the region (A.A)
This is not co-ordinates-Need to find the proper co-ordinates
```{r map_plot, include=TRUE}
map1<- get_map(location = c(lon = 18.895099, lat = -34.362008),
                      zoom = 10, maptype = 'satellite')

ggmap(map1) +
  geom_point(data = co_ordinates, aes(x = lon+0.002, y = lat-0.007),
             colour = "red", size = 1.5) +
  labs(x = "Latitude", y = "Longitude", title = "Map showing the Cederberg region") +
  theme(plot.title = element_text(hjust = 0.5))


trial1 <- ggmap(map1) +
  geom_point(data = co_ordinates, aes(x = lon+0.002, y = lat-0.007),
             colour = "red", size = 1.5) +
  labs(x = "Lat", y = "Lon", title = "Map showing the Cederberg region") +
  theme(plot.title = element_text(hjust = 0.5))



new1 <- trial1 +
  geom_text(data = co_ordinates, aes(lon+0.0135, lat-0.0045, label = Site),
            hjust = 1.05, vjust = 0.7,
            size = 3, colour = "white")
```


Some very basic analysis to get the project started (A.A)


```{r}
dat1 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.pH = mean(pH),
            sd.pH = sd(pH))

plot_A <- ggplot(dat1, aes(x = Site, y = mn.pH)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.pH - sd.pH, 
                    ymax = mn.pH + sd.pH)) +
  labs(x = "Site", y = "Average pH") +
  theme_AJMS()

dat2 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Salinity = mean(Salinity),
            sd.Salinity = sd(Salinity))

plot_B <- ggplot(dat2, aes(x = Site, y = mn.Salinity)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Salinity - sd.Salinity, 
                    ymax = mn.Salinity + sd.Salinity)) +
  labs(x = "Site", y = "Average Salinity") +
  theme_AJMS()

dat3 <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(mn.Temperature = mean(Temperature),
            sd.Temperature = sd(Temperature))

plot_C <- ggplot(dat3, aes(x = Site, y = mn.Temperature)) +
  geom_col(aes(fill = Site)) +
  geom_errorbar(aes(ymin = mn.Temperature- sd.Temperature, 
                    ymax = mn.Temperature + sd.Temperature)) +
  labs(x = "Site", y = "Average Temperature") +
  theme_AJMS()


Final <- ggarrange(plot_A, plot_B,plot_C,
                   ncol = 3, nrow = 1,
                   common.legend = TRUE)
Final
```

```{r anova1, include=TRUE}
Anova1 <- aov(pH~Site,data = R_assignment_AJ_)
summary(Anova1)
```

```{r anova2, include=TRUE}
Anova1 <- aov(Temperature~Site,data = R_assignment_AJ_)
summary(Anova1)
```


Does temperature affect the number of rhizomes present
```{r}
plot_1.2 <- ggplot(data = R_assignment_AJ_, aes(x = pH, y = rhizomes)) + 
  geom_col(aes(fill = Site)) +
  theme_AJMS()

plot_1.3 <- ggplot(data = R_assignment_AJ_, aes(x = Temperature, y = rhizomes)) + 
  geom_col(aes(fill = Site)) +
  theme_AJMS()

plot_1.4<- ggplot(data = R_assignment_AJ_, aes(x = Depth, y = rhizomes)) + 
  geom_col(aes(fill = Site)) +
  theme_AJMS()

final1 <- ggarrange(plot_1.2, plot_1.3,plot_1.4,
                   ncol = 3, nrow = 1,
                   common.legend = TRUE)
final1
```

```{r}
summary(R_assignment_AJ_$Temperature)
data <- R_assignment_AJ_
data %>%
  summarise(min_t = min(Temperature),
            quart_1 = quantile(Temperature, 0.25),
            med_t = median(Temperature),
            mean_t = mean(Temperature),
            quart_3 = quantile(Temperature, 0.75),
            max_t = max(Temperature))

ggplot(data, aes(x = Depth, y = Temperature)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_AJMS()
```


```{r}
# Calculate sumarry stats for plotting over the box plots

summary_stats <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(temp_mean = mean(Temperature))

# Plot these
plot_1 <- ggplot(data = R_assignment_AJ_, aes(x = Site, y = Temperature)) +
  geom_boxplot(aes(fill = Site), notch = TRUE) +
  geom_point(data = summary_stats,  size = 6, shape = 18,
             aes(y = temp_mean, colour = "Green")) +
  theme_AJMS()


# Calculate sumarry stats for plotting over the box plots

summary_stats <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(pH_mean = mean(pH))

# Plot these
plot_2 <- ggplot(data = R_assignment_AJ_, aes(x = Site, y = pH)) +
  geom_boxplot(aes(fill = Site), notch = TRUE) +
  geom_point(data = summary_stats,  size = 6, shape = 18,
             aes(y = pH_mean, colour = "Green")) +
  theme_AJMS()

# Calculate sumarry stats for plotting over the box plots

summary_stats <- R_assignment_AJ_ %>% 
  group_by(Site) %>% 
  summarise(rhizomes_mean = mean(rhizomes))

# Plot these
plot_3 <- ggplot(data = R_assignment_AJ_, aes(x = Site, y = rhizomes)) +
  geom_boxplot(aes(fill = Site), notch = TRUE) +
  geom_point(data = summary_stats,  size = 6, shape = 18,
             aes(y = rhizomes_mean, colour = "Green")) +
  theme_AJMS()


Final <- ggarrange(plot_1, plot_2,plot_3,
                   ncol = 3, nrow = 1,
                   common.legend = TRUE)
Final

```

```{r}
# Areas and adding trend
ggplot(data = R_assignment_AJ_, aes(y = rhizomes, x = Depth)) +
  geom_point(aes(colour = Site)) +
  geom_smooth(aes(colour = Site), method = "lm") +
  coord_equal(xlim = c(0, 60), ylim = c(0,60))
# Gray shade is the standard error
# angle of line shows the relationship
```


```{r pearsons1, include=TRUE}
# Pearsons correlations
cor.test(R_assignment_AJ_$Temperature, R_assignment_AJ_$Depth, method = "pearson",
         conf.level = 0.95)

```

```{r pearsons2, include=TRUE}
# Pearsons correlations
cor.test(R_assignment_AJ_$rhizomes, R_assignment_AJ_$Depth, method = "pearson",
         conf.level = 0.95)

```


compaing temperature/pH/rhizomes to depth
Test if data is normally distributed or not. If data is not normaly distibuted then use this :
for a t-test we rather will use Wilcox rank sum test
wilcox.test() # this is then filled with the same as for the t-test
Things to do (A.A): Group input....
Check for normality
do different stats tests on variables
what other variabbles should we include and why?
What should be excluded and why?
Any different analysis that could be added into the task that was not done before? WHY?
Statistical analysis....







